Random Matrix Theory Tools for the Predictive Analysis of Functional Magnetic Resonance Imaging Examinations

### 2023-04-07 paper/paper/SPIE/article.tex
WIP

Simplifying somewhat, a random matrix is a matrix that has entries that can be
well-described as being drawn from a particular independent and identically
distributed (\textit{iid}) distribution, or, equivalently, is a matrix where
each entry is itself a random variable. A \textit{class} or \textit{ensemble}
of random matrices is generally defined by the shape of the \textit{iid}
distribution involved. For example, the Gaussian Orthogonal Ensemble (GOE) is a
\textit{class} or \textit{ensemble} of random matrices that comprises
orthogonal matrices with each entry being sampled from a standard Gaussian, and
the Marchenko-Pastur distribution describes the limiting behaviour of the
singular values of \textit{iid} rectangular random matrices where the
\textit{iid} distribution is arbitrary\cite{mehtaRandomMatrices2004}. Of
course, given a single particular matrix, one can only estimate a
\textiit{probability} that it is a member of a certain class, since certain RMT
classes can yield arbitrary matrices (albeit with extremely low probability).

@@

Eigenvalue-based features inspired by RMT, and extracted from fMRI functional
connectivity matrices were found to have predictive utility across a wide
variety of datasets and classification tasks. However, the predictive utilities
were modest, and highly dependent on preprocessing steps and other fitting and
feature selection procedures. Future studies employing RMT should carefully
investigate the sensitivity of any findings to such analytic decisions before
drawing any conclusions. However, this study demonstrates that the inclusion of
RMT statistics and other eigenfeatures extracted from fMRI images could aid in
the predictive modeling of a wide variety of phenomena.

>>>>

Simplifying somewhat, a random matrix is a matrix that has entries that can be
well-described as being drawn from a particular independent and identically
distributed (\textit{iid}) distribution, or, equivalently, is a matrix where
each entry is itself a random variable. A \textit{class} or \textit{ensemble}
of random matrices is generally defined by the shape of the \textit{iid}
distribution involved. For example, the Gaussian Orthogonal Ensemble (GOE) is a
\textit{class} or \textit{ensemble} of random matrices that comprises
orthogonal matrices with each entry being sampled from a standard Gaussian, and
the Marchenko-Pastur distribution describes the limiting behaviour of the
singular values of \textit{iid} rectangular random matrices where the
\textit{iid} distribution is arbitrary\cite{mehtaRandomMatrices2004}. Of
course, given a single particular matrix, one can only estimate a
\textit{probability} that it is a member of a certain class, since certain RMT
classes can yield arbitrary matrices (albeit with extremely low probability).

@@

Eigenvalue-based features inspired by RMT, and extracted from fMRI functional
connectivity matrices were found to have predictive utility across a wide
variety of datasets and classification tasks. However, the predictive utilities
were modest, and highly dependent on preprocessing steps and other fitting and
feature selection procedures.

\color{cyan}
Previous studies have not explored the sensitivity of RMT-based analyses to
analytic decisions. It is far too early to say whether RMT may prove fruitful
for understanding fMRI, however, we hope our analyses show that, \textit{if}
RMT has such potential, that potential must be contextualized by the analytic
decisions involved. Absent this context, we advise caution with regard to
claims that RMT offers insights into fMRI analyses.

Future studies employing RMT should carefully
investigate the sensitivity of any findings to such analytic decisions before
drawing any conclusions. However, this study demonstrates that the inclusion of
RMT statistics and other eigenfeatures extracted from fMRI images could aid in
the predictive modeling of a wide variety of phenomena.
